---
title: "NOAA Package Overview"
author: "Chris Taber"
date: "2019-08-09"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NOAA Package Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(knitr)
library(dplyr)
library(ggplot2)
library(NOAA)
```

This Package includes functions used to read in and map earthquake data from National Oceanic and Atmospheric Administration (NOAA), which is a American scientific agency within the United States Department of Commerce that focuses on the conditions of the oceans, major waterways, and the atmosphere. 

## Functions

**eq_get_data** This function takes a country name in all capitals, a start date as a string in 'YYYY-MM-DD' format, and optionally an end date in the same format, read in the earthquake data, cleans it, and turns a filtered cleaned dataframe.

**eq_clean_data** This function takes a dataframe and drops date related NAs, converts data types to appropriate type, removes pre-BC dates due to low information, fixes date parts that do not fit conventionall format, and returns a cleaned dataframe.

**eq_location_clean** This funtion cleans the LOCATION_NAME column by stripping out the country name (including the colon) and converts names to title case (as opposed to all caps).

**geom_timeline** This function uses ggplot to creat a custom geom to visualize a timeline of earthquakes.

**geom_timeline_label** This function creates a new geom that  can build on geom_timeline and add labels to the top n biggest earthqakes by magnitude.

**eq_map** This function plots earthquakes and avaialble information about said earthquakes on an interactive map that will allow the used to click on an earthquake to see the relevant details.

**eq_create_label** This function add date, location, magnitude, and total deaths, in an annotation on an interactive map.


**Example of *eq_get_data*:**
```{r echo=TRUE}
filename <- system.file('extdata', 'earthquakes.tsv.gz', package = 'NOAA')
data<-eq_get_data(filename, country = c('MEXICO','USA'), start_date = '2006-01-01', end_date = '2009-01-01')
head(data)
```

**Example of *geom_timeline* and *geom_timeline_label*:**
```{r echo=TRUE}
data %>% 
  ggplot2::ggplot() +
  geom_timeline(
    aes(
      x = DATE,
      y = COUNTRY,
      fill = TOTAL_DEATHS,
      size = EQ_PRIMARY)
  ) +
  geom_timeline_label(
    aes(
      x = DATE,
      y = COUNTRY,
      magnitude = EQ_PRIMARY,
      label = LOCATION_NAME,
      n_max=3) # number of lables to show

  )
```

**Example of *eq_map*:**
```{r echo=TRUE}
data %>%
     dplyr::mutate(popup_text = eq_create_label(.)) %>%
     eq_map(annot_col = "popup_text")
```


